\large{\bf Large effects and the infinitesimal model }

### 2023-11-15 review-responses.tex

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
  {\it Theoretical Population Biology}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Todd Parsons \\
    \emph{and} Peter Ralph \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revision of our manuscript, ETCETERA

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Todd Parsons and
Peter Ralph
}\\
\end{minipage}

\vskip 4em

\pagebreak

%%%%%%%%%%%%%%
\section*{AE comments:}

\begin{quote}
The two reviewers find the results interesting and worth being eventually published in the special issue. However, they highlight a number of places in the text where the properties claimed or the precise questions addressed need serious clarification. As the readership of the special issue will be quite broad, it is important that the main messages are accessible to colleagues who do not necessarily know Fisher's infinitesimal model precisely. Therefore, although the mathematical results in themselves do not require deep modifications, I recommend that a major revision of the paper should be provided, which would take into account the different comments and suggestions of the reviewers.
\end{quote}

XXX

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
The paper explores inheritance of an additive trait influenced by a large number of loci with effect sizes drawn from a power law distribution, also fitting effect size distributions as estimated by GWAS (for several human disease traits) to estimate plausible power law exponents. The study is motivated by Barton et al's 2017 paper which shows how in the limit of a very large number of loci with effect sizes drawn from a distribution with finite variance, inheritance is described by the the infinitesimal model: in particular, that the distribution of trait values of the offspring of two parents is multivariate normal with a variance-covariance matrix that is independent of parental trait values. This paper explores whether some analogous statements can be made if the underlying effect size distribution does not have finite variance- whether, loosely speaking, some ``generalisation'' of the infinitesimal model might still apply. Below I summarise some of my concerns about the paper.
\end{quote}

Thanks for the careful reading...

\begin{point}{}
    Paragraph starting ``This suggests exploring whether stable distributions..'' \revref:
The questions as stated: e.g., whether stable distributions can ``stand in'' for a Gaussian distribution in the infinitesimal model or whether the results of Barton et al ``carry over'' when the sum of effect sizes follows a stable law are quite vague and allow for multiple interpretations. Ideally, in order to address whether Barton et al's results (do not) carry over, one would like to at least show that with a power law distribution of effect sizes, the ``Mendelian noise term'', i.e., the difference between the offspring trait value and the midparent value, is (not) independent of the midparent value in the limit of a very large number of loci. And indeed this is what the authors address
somewhat non-rigorously (and for the case of a Cauchy distribution of effect sizes) in section 5 and using simulations : see e.g., figures 4C and 4f, figures 6C and 6F. Unsurprisingly, these simulations show that with e.g., Cauchy distribution of effect sizes, the non-independence between the segregation/noise term and the midparent trait value does not hold. However, a lot of the mathematical analysis in the paper (sections 3 and 4) explores when the distribution of trait values in a population is (not) Gaussian, which is a completely different question.
\end{point}

\reply{
}

\begin{point}{\revref}
In the same paragraph: I don't think the claim that ``independence of offsprings' deviations and midparent values implies a Gaussian distribution'' (which is also stated in the abstract) is true? Surely, ``independence of offsprings' deviations and midparent values'' is just the basic statement of the infinitesimal model (say, in Barton et al 2017) and the infinitesimal model does not imply a Gaussian distribution of trait values? See also section 3.
\end{point}

\reply{
}

\begin{point}{}
    Section 2 and figure \ref{fig:example_snps}: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes. There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., Simons et al (A Population Genetic Interpretation of GWAS Findings for Human Quantitative Traits) for a detailed exploration of this. Moreover, if SNPs are not truly causal but only tag causal alleles, then weighting by allele frequencies can be even more problematic?

As an alternative, would it make sense to look at what fraction of loci have effect size difference between alternative alleles to be $>2t$ without weighing by allele frequency?
More broadly, how do we interpret the frequency-weighted vs. unweighted distributions?
\end{point}

\reply{
}

\begin{point}{}
I was also wondering about the ``goodness of fit'' of observed distributions to power laws. I am not necessarily recommending a detailed quantitative analysis, but nevertheless it may be useful to comment on whether one has power to say reject alternative distributions (e.g., a mixture of exponentials) that are often considered. If not, then perhaps the exact exponents inferred are unimportant, and what the data points to more broadly, is a broad distribution of effect sizes.
\end{point}

\reply{
}

\begin{point}{}
    Section 3, Sentence starting ``We know from Barton et al (2017)...'' \revref: As far as I understand, the statement of Barton et al is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents. I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole. It'd be helpful to distinguish more carefully between conditioned and unconditioned distributions, especially in this section.
\end{point}

\reply{
}

\begin{point}{}
    In general, I find this section rather confused: the infinitesimal model basically implies that the ``Mendelian segregation term'' $R_{M}$ (which is the difference between an individual and the midparent value $\bar{Z}_{M}$) is normally distributed with mean zero and variance that depends on the relatedness between parents but is independent of the midparent value $\bar{Z}_{M}$. However, in Proposition \ref{prop:parental_contribs}, the authors claim that $\bar{Z}_{M}$ and $R_{M}$ are jointly Gaussian (which is equivalent to saying that the trait value of an individual and the mean of its parents' trait values is jointly Gaussian), which is in general not true.
\end{point}

\reply{
}

\begin{point}{}
    I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' \revref. As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).
\end{point}

\reply{
}

\begin{point}{}
    Section 4, first paragraph \revref: I don't think the third assumption (that trait values in the population follow a Gaussian distribution) is a component of the ``infinitesimal model''. I think Turelli (2017) is also quite clear about this stating that a Gaussian distribution of trait values should only emerge for Gaussian or very weak selection on the population.
\end{point}

\reply{
}

\begin{point}{}
More generally, I found it a bit hard to follow the overall logical flow of section 4.1, especially once the authors launch into an exploration of the ``reproduction'' and ``noise'' terms (pages 9 and 10).
    It would be useful to clarify what the biological/intuitive meaning of the noise and reproduction terms at the outset. It would also be useful to state (at the beginning) what the main goal of these explorations is: (is the goal to identify when both terms ``have well-posed limits independent of the other'' as stated at \llname{limit_remark}), and also to summarise (at the end) to what extent one can do this in all generality.
\end{point}

\reply{
}

\begin{point}{}
    Last inequality on Page 13: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution? If the former, then it'd be good to say this explicitly. If the latter, then this does not quite answer the question posed at the beginning of the section about how much information knowing the trait value of an individual gives about underlying allelic effects.
\end{point}

\reply{
}

\begin{point}{}
Also, maybe worth specifying what the corresponding distribution for the largest of $M$ alleles looks like when the effect size distribution has finite variance, in order to highlight the contrast between ``well-behaved'' and heavy-tailed effect size distributions.
\end{point}

\reply{
}

\subsubsection{Minor comments:}

\begin{point}{}
    Introduction: Also worth citing Fisher, Bulmer etc. (the original references) when introducing the infinitesimal model?
\end{point}

\reply{
}

\begin{point}{\revref}
    ``Perhaps the distribution of effect sizes within each gene is Normal'': I am confused by this. Do you mean that the distribution of effect sizes across all genes (with roughly the same ``proximity'' to the trait) is normal. If yes, then the above phrasing is a bit misleading. If what is meant that is really that the effect sizes of different alleles within a gene are normally distributed, then I am not sure if this is entirely plausible: see various papers by Turelli (e.g., ``Heritable Genetic Variation via Mutation-Selection Balance: Lerch's Zeta Meets the Abdominal Bristle.''). Presumably, a normal distribution of effect sizes for a sufficiently large genomic region is a good approximation but it is unclear what is large enough: smaller or larger than a typical gene...?
\end{point}

\reply{
}

\begin{point}{\revref}
    Sentence starting ``In fact, many traits claimed...'': This phrasing here is rather cryptic and I do not understand what is being said. Maybe rephrase? Also a reference would help...
\end{point}

\reply{
}

\begin{point}{}
    Figure \ref{fig:exponent_hist}: Is there also a correlation between the number of cases and the number of SNPs? In other words, to what extent are (b) and (c) independent?
\end{point}

\reply{
}

\begin{point}{\revref}
    Sentence starting ``To see what knowing the parental trait value...'': In what sense is this the ``parental'' trait value??
\end{point}

\reply{
}

\begin{point}{}
    Last inequality on Page 13: should the subscript be $j$ instead of $i$?
\end{point}

\reply{
}

\begin{point}{}
    Simulations: Why not choose the same effect size distributions in the neutral case and the case with stabilizing selection? Alternatively, if different effect size distributions are used, it may be good to plot the distribution of (appropriately) scaled trait values. Right now the figures give the impression that more variation is maintained under stabilizing selection..
\end{point}

\reply{
}

\begin{point}{}
    Figure \ref{fig:seg_noise} caption: I find the phrasing ``(a,b) chosen to span the 5\% of midparents centered on the 10\% quantile of midparent value'' a bit confusing. Does this mean a and b are respectively the 7.5\% and 12.5\% quantile of the distribution of midparent trait values?
\end{point}

\reply{
}

\begin{point}{}
    Discussion: In the context of theory combining large and small-effect loci, the 2008 paper by Chevin and Hospital: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and the 1983 work by Lande: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
\end{point}

\reply{
}

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{quote}
The authors investigate the possibility of extending results of Barton et al (2017) on the
justification of the infinitesimal model to distributions of effect sizes with a large tail (i.e.,
polynomial decay with exponent $\alpha < 2$). This is biologically interesting and important
because there are clear indications that variation in some traits is caused by a mixture
of polygenic (infinitesimal) contributions and one or few loci of large effect. It also a
mathematically difficult topic because it leads into the realm of generalized Central Limit
Theorems, namely stable laws.

After a brief, partly rough (see detailed comments below) introduction the ms starts with
a statistical analysis of the effect size distributions of a large number of traits (all human
disease related) for which GWAS data are available. The evaluation shows that almost
all estimated tail exponents are between 1 and 2.5 with a quite flat distribution between
about 1.3 and 2 (some potential statistical artifacts are noted). However, the authors do
not discuss the influence of selection, which could be essential because the traits are disease
related.
\end{quote}

We now discuss selection XXX WHERE

\begin{quote}
In Section 3, the authors show a kind of negative result concerning the possibility of $\alpha$-
stable distributions with $\alpha < 2$ (Proposition 1): If the deviations of the offspring trait
from the midparent value (here called the Mendelian sampling term, $R$) are independent
of the midparent value $\bar Z$ (part of the core result in Barton et al (2017)), then the joint
distribution of $R$ and $\bar Z$ is Gaussian. Thus, this independence is not compatible with ‘$\alpha$-
stable laws’. In fact, as they show (by example) and discuss more generally in Sect. 4, a
single polymorphic large-effect locus causes dependence of $R$ and $\bar Z$.

In Section 4, a version of the infinitesimal model is investigated that proposes a much
stronger result than a Gaussian Mendelian sampling term independent of the parental
trait distribution, namely a Gaussian distribution of the trait (within the population)
under mutation-selection balance. Although such a model has occasionally been assumed
in the literature, it has never been justified rigorously and, presumably, is not justifiable
(on the basis of realistic assumptions – as clearly noted by Turelli’s brief review and
previous extensive work by Turelli, Barton, and others). Nevertheless, the analysis of the
authors is interesting from a theoretical point of view because they investigate whether
such deterministic mutation-selection balance is possible in a non-Gaussian model that
retains independence of the Mendelian sampling terms, and if such a model might yield a
reasonable approximation to an additive model with non-Gaussian effects.
\end{quote}

We don't actually assume a Gaussian distribution....

\begin{quote}
The authors use a simplistic Moran model to investigate these questions. However, mu-
tation doesn’t seem to be modeled explicitly. Is it mutation if a parent is replaced by an
offspring of the type described in the display equation on p 7? Maybe yes, but not neces-
sarily because of segregation and recombination contribute substantially to $R$. Doesn’t the
present assumption imply that the mutation rate is one and that mutation is modeled as a
kind of diffusion? In the latter case, a Gaussian distribution can result. With explicit mu-
tation models, where mutations occur per locus and a finite number of loci determines the
trait additively, a Gaussian distribution never results (and in no known limit; see above).
How does population size enter? Is this a deterministic Moran model? Anyway, the results
in Sect 4 are of interest because they strongly constrain the possibilities to justify such
‘Gaussian-population approximation’ by any infinitesimal-like model (essentially to mix-
tures of Cauchy distributions. The noted relation to blending inheritance is intriguing!).
My main concern is that especially this section needs are clearer description of the model,
and clear distinction between this ‘maximalist’ model (which has little theoretical sup-
port) and the model investigated by Barton et al (2017). Also a much clearer summary
and discussion of the implications of the results shown here are needed. It should not be
assumed that readers have read Turelli (2017) and even less that they have studied Barton
et al (2017) in considerable detail!
\end{quote}

We now explain stuff more...

\begin{quote}
In Section 5, it is pointed out why the Barton et al. (2017) fails in the case of ‘Cauchy
noise’ (and presumably for any other limit law with $\alpha < 2$). In the absence of giving the
estimate in App D in Barton et al, the explanation at the bottom of p 12 is difficult to
follow. Please provide their estimate for easier comparison (do not assume that readers
studied Barton et al; assume that they scanned (parts) of it).
\end{quote}

\begin{quote}
Finally, Section 6 presents an insightful comparison of the distributions of trait values
resulting from either Gaussian or Cauchy mutation-effect distributions. This seems to be
an explicit genetic model, but again, what is the mutation rate? Clearly, the mutation
rate determines the variance (or more generally variability) of the population distribution.
Please clarify the model. It also needs to be pointed out clearly that such results seem to
be at variance with all (?) explicit multilocus models that have been studied so far

In summary, this ms contains several interesting aspects and will deserve publication after
improving the presentation and clarifying a number of issues.
\end{quote}

\begin{point}{}
Intro, p 2: Not too many readers will be familiar with $\alpha$-stable distributions. It may be
useful to clearly point out the difference to the Gaussian case and what happens if $\alpha \ge 2$
(obviously, this is of relevance).
\end{point}

\reply{
}

\begin{point}{}
p 3, top: State that M is the number of loci.
\end{point}

\reply{
}

\begin{point}{\revref}
Should this read ‘between 20\% and 10\%’? Why this choice?
\end{point}

\reply{
}

\begin{point}{\revref}
First line in Sect 3: add ‘as $M \to \infty$’. In the next line, this is first time that $\alpha = 2$ is
mentioned. Clarify earlier that this case is very different from $\alpha < 2$ (maybe already below
the last display eq on p3).
\end{point}

\reply{
}

\begin{point}{\revref}
What means ‘nearby alleles’ ? Do you mean physically linked
loci? Another 3 lines below: alleles are not polymorphic; loci can be polymorphic.
\end{point}

\reply{
}

\begin{point}{}
p 7, Sect 4: As noted above, please describe this model in more detail to distinguish it
from that in Barton et al (2017), which makes much weaker assumptions. Note that it is
not justifiable except under quite extreme assumptions.
\end{point}

\reply{
}

\begin{point}{}
Third display equation in Sect 4.1 \revref: The index k in the product should start at k = 0.
\end{point}

\reply{
}

\begin{point}{}
p 8, last line: Delete ‘to’
\end{point}

\reply{
}

\begin{point}{}
p 10: It might be useful to mention earlier that the Cauchy distribution is an archetypical
example for an $\alpha$-stable distribution. The enthusiasm among biologists may be limited
though, because all the empirical trait distributions they find have mean and variance. I
don’t know if there are examples in which the variance grows (significantly) as the sample
size increases. In the last para of Sect 4.1, please specify ‘one will get convergence’. I
suppose this refers again to $M \to \infty$.
\end{point}

\reply{
}

\begin{point}{}
p 12, last display equation: Please relate this explicitly to the corresponding result in
Barton et al. (e.g., by stating their estimate).
\end{point}

\reply{
}

\begin{point}{}
p 16: This discussion is too vague. At least, adding a summary of the main findings
would be appropriate (always noting for which version of the infinitesimal model).
\end{point}

\reply{
}

\begin{point}{}
p 16, end of discussion: Does ‘Lévy stable distribution’ refer to the family of $\alpha$-stable
distributions? (I don’t think it has not been used before.)
\end{point}

\reply{
}

### 2023-11-15 writeup.tex
affils
sign error and comments
reviews put in latex

@@

\author{ \begin{small}
    Todd L. Parsons (LPSM, Sorbonne Universit\'e, CNRS UMR 8001, Paris, France, 75005) \&
    Peter Ralph
\end{small}}

@@

As suggested by its name,
under the infinitesimal model the effects of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical understanding of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many - if not most - lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps the distribution of effect sizes within each gene is Normal,
but with a standard deviation that varies with ``proximity'' to the trait;
such mixtures can easily have large proportions of extreme values.

@@

This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
in the infinitesimal model.
In this paper, we make preliminary explorations in that direction:
Do the results of \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
The immediate answer to this will be ``no'' --
in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution in some sense.
This independence is a key reason why the infinitesimal model is tractable,
and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
by ``trait-only'' models that do not explicitly represent underlying genetic variation.
Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
we then go on to analyze exactly these models.
(As motivation,
we have not shown that such models are \emph{not} good approximations to reality,
and it is interesting and perhaps informative to understand how the noise distribution affects results.)

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits claimed to be genetically determined
in fact only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race.
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Since the mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen allele at a uniformly chosen locus.
So, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail
(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
is between 20 and 10\% of all SNPs).
Examples are shown in Figure~\ref{fig:example_snps}.

@@

We know from \citet{barton2017infinitesimal} that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$,
then with $\alpha=2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.

@@

Since models can be easily set up for which trait distributions are not Gaussian,
the implication of this is that for such models, independence of the Mendelian sampling term
is not likely a good assumption.
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between nearby allelic effects.
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.

@@

As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.  The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
Leaving aside
% -- as the Gaussian-population approximation frequently does --
the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if)
deterministic mutation-selection balance is possible in a non-Gaussian model
that retains independence of the Mendelian sampling terms.
(Indeed, above we have put substantial restrictions on what models of inheritance
might lead to such models.)
We might also ask whether such a model
might still reasonably approximate an additive model with non-Gaussian effects,
and for the purpose of investigating this with simulations,
some calculations will be helpful.

@@

>>>>

\newif\ifsubmission
\submissiontrue
% \submissionfalse
\ifsubmission
    % % stuff for submission
    \usepackage{lineno}
    % allow line numbers around math environments
    % from https://tex.stackexchange.com/questions/43648/why-doesnt-lineno-number-a-paragraph-when-it-is-followed-by-an-align-equation
        \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
          \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
          \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
          \renewenvironment{#1}%
             {\linenomath\csname old#1\endcsname}%
             {\csname oldend#1\endcsname\endlinenomath}}% 
        \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
          \patchAmsMathEnvironmentForLineno{#1}%
          \patchAmsMathEnvironmentForLineno{#1*}}%
        \AtBeginDocument{%
        \patchBothAmsMathEnvironmentsForLineno{equation}%
        \patchBothAmsMathEnvironmentsForLineno{align}%
        \patchBothAmsMathEnvironmentsForLineno{flalign}%
        \patchBothAmsMathEnvironmentsForLineno{alignat}%
        \patchBothAmsMathEnvironmentsForLineno{gather}%
        \patchBothAmsMathEnvironmentsForLineno{multline}%
        }
\else
    % for the nice version
\fi

% Responses to reviews
\input{review-response-commands}
% set this to show line numbers and include responses to reviews or not
\newif\ifreviewresponses
\reviewresponsestrue  % include them
% \reviewresponsesfalse  % don't include them
\newcommand{\responsefile}{review-responses}  % name of the review responses file

@@

\author{ \begin{small}
    Todd L. Parsons\footnote{LPSM, Sorbonne Universit\'e, CNRS UMR 8001, Paris, France, 75005}
    ~\&~
    Peter L. Ralph\footnote{Institute of Ecology \& Evolution, University of Oregon, Eugene, OR, USA, 97405}
\end{small}}

@@

As suggested by its name,
under the infinitesimal model the effects of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical understanding of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many - if not most - lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps the distribution of effect sizes within each gene is Normal, \revpoint{1}{13}
but with a standard deviation that varies with ``proximity'' to the trait;
such mixtures can easily have large proportions of extreme values.

@@

This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
in the infinitesimal model. \revpoint{1}{1} \revpoint{1}{2}
In this paper, we make preliminary explorations in that direction:
Do the results of \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
The immediate answer to this will be ``no'' --
in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution in some sense.
This independence is a key reason why the infinitesimal model is tractable,
and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
by ``trait-only'' models that do not explicitly represent underlying genetic variation.
Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
we then go on to analyze exactly these models.
(As motivation,
we have not shown that such models are \emph{not} good approximations to reality,
and it is interesting and perhaps informative to understand how the noise distribution affects results.)

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits claimed to be genetically determined \revpoint{1}{14}
in fact only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race.
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Since the mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen allele at a uniformly chosen locus.
So, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail
(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
is between 20 and 10\% of all SNPs). \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$,
then with $\alpha=2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.

@@

Since models can be easily set up for which trait distributions are not Gaussian,
the implication of this is that for such models, independence of the Mendelian sampling term
is not likely a good assumption. \revpoint{1}{7}
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between nearby allelic effects. \revpoint{2}{5}
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.

@@

As observed in \cite{turelli2017commentary}, \revpoint{1}{8}
there are at least three interpretations of what is called ``the infinitesimal model''.  The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
Leaving aside
% -- as the Gaussian-population approximation frequently does --
the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if)
deterministic mutation-selection balance is possible in a non-Gaussian model
that retains independence of the Mendelian sampling terms.
(Indeed, above we have put substantial restrictions on what models of inheritance
might lead to such models.)
We might also ask whether such a model
might still reasonably approximate an additive model with non-Gaussian effects,
and for the purpose of investigating this with simulations,
some calculations will be helpful.

@@

% reviews here if toggled above
% \ifsubmission\processdelayedfloats\fi

\includereviews
